---
title: "Marine Community Ecology 2024"
subtitle: "01-Introduction to coding in R"
author: "Simon J. Brandl"
institute: "The University of Texas at Austin"
date: "2024/01/01 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    toc: yes
    css: ["xaringan-themer.css"]
    lib_dir: libs
    nature:
      highlightStyle: github
      ratio: 16:9
      highlightLines: true
      countIncrementalSlides: false
---
```{r xaringan-themer, include=FALSE, warning=FALSE}
### setup chunk
# load packages
library(xaringanthemer)
library(fontawesome)
library(emo)
library(tidyverse)
style_duo_accent_inverse(
  primary_color = "#85D3D1",
  secondary_color = "#0D2A47",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```

# Load package 

Load package for class exercises:

- Package: tidyverse
```{r, echo = TRUE}
library(tidyverse)
```

# Exercise 1 
Create the following vectors

1) a sequence of even numbers from 2 to 20
2) 1 to 9 twice
3) 2 2 2 3 3 3 4 4 4 5 5 5 6 6 6 7 7 7
4) the word "fish" 18 times
5) the words "blue", "white", and "yellow" six times

```{r echo = TRUE}
# 1) a sequence of even numbers from 2 to 20
seq(2, 20, 2)

# 2) 1 to 9 twice
rep(1:9, 2) 

# 3) 2 2 2 3 3 3 4 4 4 5 5 5 6 6 6 7 7 7
rep(2:7, each = 3)

# 4) the word "fish" 18 times
rep("fish", 18)

# 5) the words "blue", "white", and "yellow" six times
rep(c("blue", "white", "yellow"), each = 6)
```

# Exercise 2 
1) create a vector with random integer values using rpois(100, 10)
2) create a second vector with values that are smaller or equal to 10 but not 5
3) sort the second vector in ascending order 
4) calculate the mean and standard deviations of the second vector 
5) calculate the maximum value of the second vector

```{r echo = TRUE}
# 1) create a vector with random integer values using rpois(100, 10)
vector <- rpois(100, 10)
vector

# 2) calculate the number of values that are smaller or equal to 10 but not 5
vector2 <- vector[vector <= 10 & vector != 5]
vector2

# 3) sort the vector in ascending order 
sort(vector2)

# 4) calculate the mean and standard deviations 
mean(vector2)
sd(vector2)

# 5) calculate the minimum and maximum values
max(vector2)
min(vector2)
```

# Exercise 3
1) create your own matrix
2) convert it to a data frame
3) give your data frame row and column names
4) convert it to a tibble

```{r, echo=TRUE}
# 1) create your own matrix
mx <- matrix(rpois(30, 5), nrow = 5)
mx

# 2) convert it to a data frame
df <- as.data.frame(mx)
df

# 3) give your data frame row and column names
colnames(df) <- c("african.penguin", "little.penguin", "rockhopper.penguin", "emperor.penguin", "chinstrap.penguin", "adelie.penguin")
rownames(df) <- c("site1", "site2", "site3", "site4", "site5")
df

# 4) convert it to a tibble
tb <- as_tibble(df)
tb
```

# Exercise 4
1) download the dataset called "coralreefherbivores.csv" from Canvas, and put it in your working directory in a folder called "data"
2) read in the data using the read.csv() function, specifying correct path to the "data" folder
3) calculate the mean of the column "snoutlength"

```{r, echo=TRUE}
# 1) read in the data using the read.csv() function, specifying correct path to the "data" folder
data <- read.csv(file = "data/coralreefherbivores.csv")

# 2) calculate the mean of the column "snoutlength"
mean(data$snoutlength)
```